\bf Title

### 2023-10-25 article/paper_padic_spec.tex
spectrahedron and polyhedron

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{algorithm}{Algorithm}
\newtheorem{conjecture}{Conjecture}
\newtheorem{condition}{Condition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}
\usepackage{booktabs}

@@

\section{Polyhedra and linear programming}

\subsection{Polyhedra}
\subsection{Linear programmming}

@@

\subsection{Definition}
\subsection{Annuli (couronnes)}
\subsection{Polyhedrality}
Cf. \cite{bhardwaj2015deciding}

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{corollary}
\newtheorem{algorithm}{Algorithm}
\newtheorem{conjecture}{Conjecture}
\newtheorem{condition}{Condition}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}
\usepackage{booktabs}

@@

\section{Polyhedra and linear programming}
\subsection{Polyhedra}

\begin{definition}
	We call {\it convex $p$-adic polyhedra} subsets of $K^s$ defined by linear inequalities $\val\left( \ell_1(x)\right)\ge 0$,\ldots, $ \val\left( \ell_n\left( x \right) \right)\ge 0$ for $ \ell_1,\ldots, \ell_n $ fixed linear forms and for all $x \in K ^s$.

\end{definition}

\begin{example}
	The closed unit ball of $K^s$ for the infinity norm [not sure if it is really used ] is convex polyhedron.It is the set of $\left( x_1,\ldots,x_s \right) \in  ^s$ so that $$
	\begin{pmatrix} x_1 &  & & \\
		& x_2 &0 &  \\
		&  0& \ddots & \\
		&  & & x_s/
	\end{pmatrix} \ge 0.$$ which is in fact $\OK^s$.

\end{example}

\subsection{linear programming}
	We call { \it $p$-adic programming} the problem :
\begin{equation}
	\tag{PLp}
	\begin{matrix}
		\text{Minimize } \val\left(\left<c,x \right>\right) \text{ subject to }\\
		Ax + b \ge 0
	\end{matrix}
	\label{eqn:Proglinp}
\end{equation}

with $x$ a vector of $K ^n$, and $c$ a fixed vector of size $n$ that represents the cost, $A$ a fixed $m \times n $ matrix and $b$ fiexed a vector of size $m$.

The method used in this article consists in reducing the Smith normal form of the matrix $A$.

\begin{definition} Smith normal form

	We call {\it Smith normal form} of a matrix $M \in \mathcal{M}_{m,n}\left(K \right) $ of rank $r$ the only matrix $S$ of the form $$S =  
	\begin{pmatrix} p^{a_1} & \\
		& \ddots \\
		& & p^{ar}\\
		& & & 0\\
		& & & & \ddots \end{pmatrix} $$
	so that $(a_1)_{1\le i\le r}$ is a increasing sequence of integers and $M$ can be written $M =  Q^{-1} S P$ for $P \in \mathcal{G}L_n\left( \OK \right) $, $Q \in \mathcal{G}L_m\left( \OK \right) $.
\end{definition}

\begin{remark}
	\begin{enumerate}%[label=\roman*.]
		\item The coefficient of the Smith normal Form are unique and called \textit{invariant factors}.
		\item The $p$-adic valuation of the first coefficient of the Smith normal form of a matrix $M \in \mathcal{M}_n \left( K \right) $ is the minimum of the valuations of the matrix $M$'s terms.
		\item In particular, the the normal Smith form of matrix of $\mathcal{M}_n \left( \OK \right) $ have coefficients in $\OK$ .
		\item The $r = \text{rank} M$ first diagonals coefficients of $S$ are exactly its non-zero coefficient.
	\end{enumerate}

\end{remark}

In order to be able to use the Smith normal form to resolve \ref{eqn:Proglinp} we need the following lemma :

\begin{lemma}
	Forall $p$-adic numbers $z \in K ^n$ and matrix $M \in \mathcal{M}_{m,n}\left( K  \right)  $ if $z\ge 0$ and $M\ge 0$ then $Mz\ge 0$.  
\end{lemma}
\begin{proof}
	$\OK$ is a ring.
\end{proof}

If we put the matrix $A$ of \ref{eqn:Proglinp} in its normal Smith form, it comes that resolving \ref{eqn:Proglinp} is equivalent to resolve :

\begin{equation}
	\tag{PLp'}
	\begin{matrix}
		\text{Minimize } \val\left(\left<c',x \right>\right) \text{ subject to }\\
		Sy + b' \ge 0
	\end{matrix}
	\label{eqn:Proglinp2}
\end{equation}
where $b' = Qb$, $c' = P^Tc$, $S$ is the Smith normal form of $A$ and $A = Q^{-1} S P$ for some $P \in \mathcal{G}L_m\left( \OK \right)$ and $ Q \in \mathcal{G}L_n\left( \OK \right)$.

\begin{remark}

	We have then the following results
	\begin{enumerate}
		\item $x^*$ is a feasible solution of \ref{eqn:Proglinp} if and only if $y^* = P x^*$ is a feasible solution of \ref{eqn:Proglinp2}
		\item \ref{eqn:Proglinp2} have feasible solutions if and only if the $m-r$ coefficients of $b'$ are non-zero, with $r$ the rank of $S$.
		\item If at least one of the $n-r$ last coefficients of $c'$ is non-zero \ref{eqn:Proglinp2} is unbounded and does not have a soution.
	\end{enumerate}

\end{remark}

\begin{proposition}
	\begin{itemize}~

		\item[$\circ$] If not all the $m-r$ coefficients of $b'$ are zeros then there is no $y \in K ^n$ so that $A'y+b' \ge  0$.
		\item[$\circ$] If not all the $n-r$ coefficients of $c'$ are zero then the problem is unbounded and does not have solution. \end{itemize} 
\end{proposition}
If one considers only the itération of the problem which have des solution one can reduce the problem by only taking acount the $r$ first coefficients of $y, b', c'$ and the submatrix of $S$ made with only $r$ first lines et columns of and which coefficients are exactly the nonzero invariants factors. The set $Adm$ of the feasible solutions can then be written as the set of vectors $y \in \mathbb{Q}_{ p } ^n$ that verifies $\forall 1 \le i\le r \ s_i y_i + b'_i \in \OK$ i.e. that verifies :

\begin{equation}
	\forall 1 \le i\le r  \ y_i \in -\frac{b’_i}{s_i} + \frac{1}{s_i} \OK
\end{equation}

Resolving \ref{eqn:Proglinp2} consists in minimizing $\val\left(\left<c',y \right>\right)$ over $Adm$. The image of $Adm$ by $y \mapsto \left<c',y \right>$ is $\sum_{i=1}^r -c'_i.\frac{b'_i}{s_i} + \sum_{i=1}^r\left( \frac{c'_i}{s_{i}} \OK \right)$ which can be rewritten in 
$$\left<c',Adm \right> = \lambda + p^{v} \OK$$
with $\lambda = \sum\limits_{i=1}^r -c'_i.\frac{b'_i}{s_i}$ et $v = \min\limits_{1\le i\le r} \val \frac{c'_{i}}{s_{i}} $.

The result can be split depending on the valuation of $\lambda$.
\begin{itemize}
	\item If $\val( \lambda) < v$ the minimum of $y\mapsto \val\left(\left<c',y \right>\right)$ over $Adm$ is reached in any point of $Adm$ and is equal to $\val\left( \lambda\right)$.
	\item If $\val\left( \lambda \right) \ge v$ then $\lambda \in p^{v} \OK$ and the minimum is $v $ and is reached in any points $y$ of $Adm$ that verifies 
	$$\val\sum \limits_{\begin{array}{c} 1\le i\le r\\ \val\left(c'_{i}/{s_{i}} \right) = v  \end{array}   } y_{i} = 0.$$

\end{itemize}
\begin{remark}
	To maximize the valuation of a linear map over a convex $p$-adic polyhedron instead of minimizing it (which is equuivalent to minimize the absolute value) the same reasoning can be apllied. The only differnece being that if $\val\left( \lambda\right) < v$ the problem is unbounded and then does not have any solution. 
\end{remark}

@@

\subsection{Definition}
\newcommand\mat{postive semidefinite matrix } 
\newcommand\Mat{Positive semidefinite matrix }
\newcommand\mats{positive semidefinite matrices }
\newcommand\Mats{positive semidefinite matrices }
\begin{definition}
	We call {\it \mat} any matrix $M \in \mathcal{M}_n\left( K \right) $ whose eigenvalues have nonegative valuation (in $\overline{K}$) .

	We will denote as $ \mathcal{M}_n^+\left( K \right)$ the set of \mats and we will write $M \succeq 0$ if $M \in \mathcal{M}_n^+\left( K\right) $.
\end{definition}
\begin{theorem}
	\label{caracsdp}
	Characterization of the \mats

	A matrixe is positive semidefinite if and only if its characteristic polynomial have coefficients in $\OK$ .

\end{theorem}

\begin{corollary}
	$\mathcal{M}_n(\OK) \subset \mathcal{M}_n^+\left( K \right)$ 
\end{corollary}

\textit{Preuve : }  $\OK$ is a ring therefore the characteristic polynomial of matrix whose coefficients are $p$-adic integers have coefficients ins $Pp$. We conclude by \ref{caracsdp} .

\begin{remark}
	In general, the reciprocal inclusion is false [not sure about this one]. For instance the characteristic polynomial of $M = \begin{pmatrix} 5 + \frac{3}{5} & \frac{4}{5} \\ \frac{4}{5} & -\frac{3}{5} \end{pmatrix} $ is $\chi_M = X^2  - 5 X - 4$. However when seen as an element of $\mathbb{Q}_5$, $\chi_M \in \mathbb{Z}_5[X]$ then $M \in \mathcal{M}_2^+\left( \mathbb{Q}_5 \right)$ but no coefficient of $M$ is in $\mathbb{Z}_5 $
\end{remark}

\begin{proposition}
	The set $\mathcal{M}_n^+\left( K \right)$ is :
	\begin{enumerate}%[label = \textit{\roman*}.]
		\item open
		\item closed
		%\item a convex cone
	\end{enumerate}
\end{proposition}

\textit{Preuve :}  
Firstly, $\OK[X]$ is closed and open in $\mathbb{Q}_{ p }[X] $ equiped with the infinitiy norm $\|.\|_\infty : P = \sum_{k=1}^{n} a_k X^k \to \sup |a_k|_p$. $\OK$ is indeed the open unit ball for $\|\cdot \|_\infty$ which defines a discrete distance for which open balls are also closed.
Furthermore, $\mathcal{M}^+_{n}\left(K \right)  = \chi^{-1}( \OK[X]) $ where $\chi$ is the function that maps its characteristic polynomial to a matrix which is continuous.

%Le $iii$. est laissé en exercice au lecteur. 
% Je ne parle plus de convexité dans le rapport mais on ne sait jamais
%\input{convexemple.tex}

\begin{definition}
	We call {\it $p$-adic spectrahedron} the intersection of $\mathcal{M}_n^+\left( K \right) $ with an affine plan $\mathcal{L}$ of $\mathcal{M}_n\left( K \right) $.
\end{definition}

\begin{remark}
	As in the real case, the spectrahedron associated with the affine plan $\mathcal{L}$ will often be identified with its preimage in $K^s$ (i.e. the set of vectors $x \in K ^s$ so that the linear matrix $A(x) = A_0 + x_1A_1 + \ldots + x_sA_s$ is positive semidefinite with $A_0,A_1,\ldots,A_s$ spanning $\mathcal{L}$). 
\end{remark}

The latter allows gives us the following property :
\begin{proposition}
	A convex $p$-adic polyhedron is $p$-adic spectrahedron.
\end{proposition}
\begin{proof}
	Same as in the real case (maybe modify it or add the link to real proof).
\end{proof}

\subsection{Annuli (couronnes)}
\begin{definition}
	Annulus

	We call {\it annulus} a set $A$ defined as $C = \{x \in K  | a\le \val\left(x\right) \le b\} $ for $a < b$ two nonnegative real numbers.
\end{definition}

\begin{theorem}
	Annuli are spectrahedral shadows.
\end{theorem}

\textit{Preuve :}
Let $A$ be the annuli defined by $a<b \in R^*_+$.

We consider for all $x,y \in K $ the matrix  $$M(x,y) :=
\begin{pmatrix} 
	p^ax & 0 & 0 & 0 \\
	0 & p^{-b}y & 0 & 0 \\
	0 & 0 & p^{-1} & p^{-1}x \\
	0 & 0 & p^{-1}y & - p^{-1}  \\
\end{pmatrix} $$ and the associated spectrahedron $\mathcal{S}= \{(x,y) \in K  : M(x,y) \succeq 0\} $

Let us show that $x \in C$ if and only if  $\exists y \in K \left( x,y \right) \in \mathcal{S}$.

Let $x,y$ be two $p$-adic numbers.
$M(x,y)$ can be decomposed in three blocks : $p^ax$, $p^{-b}y $ and $M'(x,y) = \begin{pmatrix} p^{-1} & p^{-1}x \\ p^{-1} y & p^{-1}\end{pmatrix} $. Using \ref{caracsdp} we know that $M(x,y)$ is positive semidefinfite if and only if
$
\begin{cases}
	p^ax \ge 0 \\
	p^{-b}y \ge 0\\
	\text{Tr}M'\left( x,y \right) = 0 \ge 0\\
	\text{det} M'(x,y ) = p^{-2}\left( xy-1 \right)  \ge 0 
\end{cases}
$ 
that is if and only if 
\begin{equation}
	\label{eq:caracsdp} 
	\begin{cases} 
		\val\left(x\right)\ge a\\
		\val\left(y\right)\ge -b\\
		p^{-2} \left( xy-1 \right) \ge 0
	\end{cases}
\end{equation}

Therefore $x \in C$ if and only if there exists $y \in K $ so that $(x, y)$ verifies \ref{eq:caracsdp}. Si $x, y$ verifies indeed \ref{caracsdp} then $\val\left(x\right)\ge a$, $\val\left(y\right)\ge -b$ and $p^{-2} \left( xy-1 \right) \ge 0$ implies that $\val\left(x\right)+\val\left(y\right)=\val\left(xy\right) = \val\left(-1\right) =0$ and therefore that $\val\left(x\right)=-\val\left(y\right)\le -(-b)  =b$, so $x \in $. Reciprocally if $x \in C$ then $(x,x^{-1}) $ verifies \ref{eq:caracsdp}  .

\subsection{Polyhedrality}
Cf. \cite{bhardwaj2015deciding}
